\section{Injection Vulnerability}

\begin{frame}{Injection: same old story}

\begin{block}{SQL injection}
  \begin{itemize}
    \item Introduced by sloppy \emph{(PHP)} programmers.
    \item One of the main vulnerability in 2000s.
    \item Still present and exploited!
  \end{itemize}
\end{block}

\begin{block}{Shellshock vulnerability}
  \begin{itemize}
    \item Discovered in 2014.
    \item Present since September 1989!
    \item In some scenarios, it permits remote code execution.
  \end{itemize}
\end{block}
\end{frame}

  \lstset{basicstyle=\small, frame=single, breaklines=true, escapeinside={\%*}{*} }
\begin{frame}[fragile]{SQLi example}
  \begin{itemize}
    \item \textbf{Scenario}: system containing secret stuff on the database, which can be accessed only by knowing their secret identifier.
    \item \textbf{Objective}: obtain all the secrets!
    \item SQL Query:
\ccode
\begin{lstlisting}
exec("SELECT * FROM Secrets WHERE SecId = " + secretId)
\end{lstlisting}
\pause
    \item If secretId is \textbf{1234567}
\begin{lstlisting}
exec("SELECT * FROM Secrets WHERE SecId = 1234567")
\end{lstlisting}
  \end{itemize}

\end{frame}
\begin{frame}[fragile]{SQLi example}

\ccode
\begin{itemize}
  \item Query:
  \begin{lstlisting}
  exec("SELECT * FROM Secrets WHERE ID = " + secretId)
  \end{lstlisting}
\pause
  \item Injection if secretId is \textbf{0 OR 1=1}:
\begin{lstlisting}
exec("SELECT * FROM Secrets WHERE ID = 0 OR 1=1")
\end{lstlisting}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{A shellshock example}
  \begin{itemize}
    \item Request to a CGI page:
  \begin{lstlisting}
  GET /vulnerable.cgi
  User-Agent: Mozilla Firefox
  \end{lstlisting}
\pause
    \item Response:
  \begin{lstlisting}
  200 OK

  Hi, I'm not vulnerable!
  \end{lstlisting}
  \end{itemize}

\end{frame}

\begin{frame}[fragile]{A shellshock example}

  \begin{itemize}
    \item Request:
  \begin{lstlisting}
  GET /vulnerable.cgi
  User-Agent: () { ignored; }; echo The following sentence is false
  \end{lstlisting}
\pause
    \item Response:
  \begin{lstlisting}
  200 OK
  The following sentence is false

  Hi, I'm not vulnerable!
  \end{lstlisting}
  \end{itemize}
\end{frame}

\begin{frame}{Data and code are confused}
\begin{block}{Untrusted user-supplied inputs}
  \begin{itemize}
    \item Web form fields (yes, even hidden tags).
    \item HTTP fields.
    \item Shell parameters.
  \end{itemize}
\end{block}

\begin{block}{No difference between data and code}
  \begin{itemize}
    \item No data type.
    \item Semantic checks are performed only on the result.
  \end{itemize}
\end{block}
\end{frame}

\begin{frame}{Sanitization}
  \begin{itemize}
    \item \textbf{Whitelist} allowed character. Drop or escape the rest.
    \item \textbf{Type awareness:} parse integers, encode strings, etc..
    \item Use \textbf{framework or library functions} considered secure.
    \item Remember: \textbf{security isn't just a fix}.
  \end{itemize}
\end{frame}
